### 2023-06-16 application.tex
section 3

\section{Computing the maximal resilient subsystem}

Let $\mathscr{S}=(S, \rightarrow, \leq)$ be WSTS and $X \subseteq S$ be an upward-closed set given by its finite minimal basis $B_X$.

\begin{definition}
We say that $\mathscr{S}=(S, \rightarrow, \leq)$ is $X$-resilient (resp. $(s_0,X)$-resilient) if $S  \xrightarrow{*} X$ (resp. $\post^*(s_0) \xrightarrow{*} X$) is satisfied.  
\end{definition}

We may verify that  $\mathscr{S}=(S, \rightarrow, \leq)$ is $X$-resilient iff  $\mathscr{S}=(S, \rightarrow, \leq)$ is $\pred^*(X)$-resilient. The \emph{maximal $X$-resilient subsystem} of $\mathscr{S}=(S,\rightarrow,\leq)$ is defined as $\mathscr{S}_X=(S,\rightarrow_X,\leq)$ where $\rightarrow_X$ is the largest restriction of $\rightarrow$ that insures that $\mathscr{S}_X$ is $X$-resilient.
%\alain{vérifier que la plus grande existe}
Let us remark that given a WSTS $\mathscr{S}=(S, \rightarrow, \leq)$, one may construct many \emph{guard} systems where every transition $s \rightarrow s'$ of $\mathscr{S}$ is guarded by a (monotone) formula $\phi$ defined by the grammar $\phi ::= s \geq u \mid s' \geq u \mid \phi \vee \phi \mid \phi \wedge \phi$ where $u \in S$. Since guards restrict the original runs but don't create new ones, every guarded transition system $\mathscr{S'}$ is a subsystem of $\mathscr{S}$. Since these formula are upward-compatible, the guarded transition systems $\mathscr{S'}=(S, \rightarrow', \leq)$ are still WSTS.
%
For example, VASS allows such guards.
  %	\lor s' \geq x_2  	\lor...	\lor s' \geq x_p$ (facile dans les VASS et modèles habituels).

\begin{theorem}{}
Given an effective WSTS with effective pred-basis and an upward-closed set $X$, the maximal resilient subsystem is a computable WSTS.
\end{theorem}

\begin{proof}
Since $X \subseteq S$ is upward-closed and $\mathscr{S}=(S, \rightarrow, \leq)$ is an effective WSTS with effective pred-basis, we compute the minimal basis of $\pred^*(X)=\{x_1,x_2,...,x_p\}$. 
%
%
Then we add the following guard $s' \geq x_1  	\vee s' \geq x_2  	\vee...	\vee s' \geq x_p$ to every transition $s \rightarrow s'$ of $\mathscr{S}$ and we define  $s \rightarrow' s'$ as $(s' \geq x_1  	\vee s' \geq x_2  	\vee...	\vee s' \geq x_p) \wedge s \rightarrow s'$ ; this insures that the new (guarded) WSTS $\mathscr{S'}=(S, \rightarrow', \leq,s_0)$ is $\pred^*(X)$-resilient and then it is also the maximal resilient sub-WSTS of $\mathscr{S}=(S, \rightarrow, \leq)$.
\end{proof}
%		l'état atteint $s'$ est au dessus des minimaux de $\pred^*(X)$, ainsi par (upward) compatibility on sera toujours capable d'aller dans $X$. 
%
%
%\noindent
%{\bf Vector addition system with states}
%
Let us recall that a {\em vector addition system with states (VASS)} in dimension $d$ ($d$-VASS for short) is a finite $\mathds{Z}^d$-labeled directed graph $V = (Q,T)$, where $Q$ is the set of {\em control-states}, and $T \subseteq Q \times \mathds{Z}^d \times Q$ is the set of {\em control-transitions}. 
% The {\em size} of $V$ is defined as $|V|=|Q|+|T|*d*|log(||T||)$ where $||T||$ denotes the absolue value of the largest number that appears in $T$, i.e. $||T|| = max\{ ||\textbf{z}||: (p,\textbf{z},q) \in T\}$.
%
Subsetquently, $Q \times \N^d$ is the set of configurations of the transition system associated with $V$.
For all configurations $p(\textbf{u}), q(\textbf{v}) \in Q \times \N^d$ and for every control-transition $t = (p, \textbf{z}, q)$ we write $p(\textbf{u}) \xrightarrow{t} q(\textbf{v})$ whenever $\textbf{v} = \textbf{u} + \textbf{z} \geq \textbf{0}$
%
\iffalse \mathieu{Defining it like this makes it a LTS rather than a unlabeled TS - maybe talk about how we can 'forget' the labels to obtain an unlabeled TS ?}
\alain{les VASS ne sont pas vraiment étiquetés, sauf si on veut le faire, car on peut dire que les VASS sont définis à partir d'un nombre fini de control-transitions  $t = (p, \textbf{z}, q)$ générant une infinité de transtions notées $p(\textbf{u}) \xrightarrow{t} q(\textbf{v})$ puisque pour tout s, $p(\textbf{u+s}) \xrightarrow{t} q(\textbf{v+s})$} \fi
When in the context of a $d$-VASS, we denote $0^d$ by $\textbf{0}$.

>>>>

\end{proof}
%		l'état atteint $s'$ est au dessus des minimaux de $\pred^*(X)$, ainsi par (upward) compatibility on sera toujours capable d'aller dans $X$. 
%
%
%\noindent
%{\bf Vector addition system with states}
%

Let us recall that a {\em vector addition system with states (VASS)} in dimension $d$ ($d$-VASS for short) is a finite $\mathds{Z}^d$-labeled directed graph $V = (Q,T)$, where $Q$ is the set of {\em control-states}, and $T \subseteq Q \times \mathds{Z}^d \times Q$ is the set of {\em control-transitions}. 
% The {\em size} of $V$ is defined as $|V|=|Q|+|T|*d*|log(||T||)$ where $||T||$ denotes the absolue value of the largest number that appears in $T$, i.e. $||T|| = max\{ ||\textbf{z}||: (p,\textbf{z},q) \in T\}$.
%
Subsetquently, $Q \times \N^d$ is the set of configurations of the transition system associated with $V$.
For all configurations $p(\textbf{u}), q(\textbf{v}) \in Q \times \N^d$ and for every control-transition $t = (p, \textbf{z}, q)$ we write $p(\textbf{u}) \xrightarrow{t} q(\textbf{v})$ whenever $\textbf{v} = \textbf{u} + \textbf{z} \geq \textbf{0}$
%
\iffalse \mathieu{Defining it like this makes it a LTS rather than a unlabeled TS - maybe talk about how we can 'forget' the labels to obtain an unlabeled TS ?}
\alain{les VASS ne sont pas vraiment étiquetés, sauf si on veut le faire, car on peut dire que les VASS sont définis à partir d'un nombre fini de control-transitions  $t = (p, \textbf{z}, q)$ générant une infinité de transtions notées $p(\textbf{u}) \xrightarrow{t} q(\textbf{v})$ puisque pour tout s, $p(\textbf{u+s}) \xrightarrow{t} q(\textbf{v+s})$} \fi
When in the context of a $d$-VASS, we denote $0^d$ by $\textbf{0}$.

### 2023-06-16 resilience.tex
section 3

>>>>

\subsection{Computing the maximal resilient subsystem}

Let $\mathscr{S}=(S, \rightarrow, \leq)$ be WSTS and $X \subseteq S$ be an upward-closed set given by its finite minimal basis $B_X$.

\begin{definition}
We say that $\mathscr{S}=(S, \rightarrow, \leq)$ is $X$-resilient (resp. $(s_0,X)$-resilient) if $S  \xrightarrow{*} X$ (resp. $\post^*(s_0) \xrightarrow{*} X$) is satisfied.  
\end{definition}

We may verify that  $\mathscr{S}=(S, \rightarrow, \leq)$ is $X$-resilient iff  $\mathscr{S}=(S, \rightarrow, \leq)$ is $\pred^*(X)$-resilient. The \emph{maximal $X$-resilient subsystem} of $\mathscr{S}=(S,\rightarrow,\leq)$ is defined as $\mathscr{S}_X=(S,\rightarrow_X,\leq)$ where $\rightarrow_X$ is the largest restriction of $\rightarrow$ that insures that $\mathscr{S}_X$ is $X$-resilient.
%\alain{vérifier que la plus grande existe}
Let us remark that given a WSTS $\mathscr{S}=(S, \rightarrow, \leq)$, one may construct many \emph{guard} systems where every transition $s \rightarrow s'$ of $\mathscr{S}$ is guarded by a (monotone) formula $\phi$ defined by the grammar $\phi ::= s \geq u \mid s' \geq u \mid \phi \vee \phi \mid \phi \wedge \phi$ where $u \in S$. Since guards restrict the original runs but don't create new ones, every guarded transition system $\mathscr{S'}$ is a subsystem of $\mathscr{S}$. Since these formula are upward-compatible, the guarded transition systems $\mathscr{S'}=(S, \rightarrow', \leq)$ are still WSTS.
%
For example, VASS allows such guards.
  %	\lor s' \geq x_2  	\lor...	\lor s' \geq x_p$ (facile dans les VASS et modèles habituels).

\begin{theorem}{}
Given an effective WSTS with effective pred-basis and an upward-closed set $X$, the maximal resilient subsystem is a computable WSTS.
\end{theorem}

\begin{proof}
Since $X \subseteq S$ is upward-closed and $\mathscr{S}=(S, \rightarrow, \leq)$ is an effective WSTS with effective pred-basis, we compute the minimal basis of $\pred^*(X)=\{x_1,x_2,...,x_p\}$. 
%
%
Then we add the following guard $s' \geq x_1  	\vee s' \geq x_2  	\vee...	\vee s' \geq x_p$ to every transition $s \rightarrow s'$ of $\mathscr{S}$ and we define  $s \rightarrow' s'$ as $(s' \geq x_1  	\vee s' \geq x_2  	\vee...	\vee s' \geq x_p) \wedge s \rightarrow s'$ ; this insures that the new (guarded) WSTS $\mathscr{S'}=(S, \rightarrow', \leq,s_0)$ is $\pred^*(X)$-resilient and then it is also the maximal resilient sub-WSTS of $\mathscr{S}=(S, \rightarrow, \leq)$.
